Exploratory data analysis

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@

```{r eda-3 }
plot_prcomp(vi_data_corrected_buffer2)
```

@@

```{r eda-6 }
plot_prcomp(nri_data_corrected_buffer2)
```

@@

```{r eda-7 }
plot_correlation(nri_data_corrected_buffer2)
```

@@

```{r eda-10 }
plot_prcomp(bands_data_corrected_buffer2)
```

@@

```{r}
map(
  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
  ~ compare_buffer(
    trees_with_bands_corrected_buffer2,
    trees_with_bands_corrected_no_buffer,
    .x
  )
) %>%
  patchwork::wrap_plots() +
  patchwork::guide_area() +
  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
  patchwork::plot_annotation(
    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
    subtitle = "Mean values of each hyperspectral band for each plot "
  )
```

>>>>

- VI = Vegetation Indices
- NRI = Normalized Ratio Indices
- HR = Hyperspectral reflectances (image bands)

@@

```{r eda-3 }
# remove response
vi_data_corrected_buffer2_no_defol_no_response <- vi_data_corrected_buffer2
vi_data_corrected_buffer2_no_defol_no_response$defoliation <- NULL

pca_vi <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol_no_response,
  variance_cap = 0.90,
  ggtheme = ggpubr::theme_pubclean(),
  geom_label_args = list(size = 3),
  parallel = TRUE
)
pca_vi[[1]]
```

@@

```{r eda-6 }
# remove response
nri_data_corrected_buffer2_no_response <- nri_data_corrected_buffer2
nri_data_corrected_buffer2_no_response$defoliation <- NULL

pca_nri <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2_no_response,
  variance_cap = 0.9,
  ggtheme = ggpubr::theme_pubclean(),
  geom_label_args = list(size = 3),
  parallel = TRUE
)
pca_nri[[1]]
```

@@

No correlation plot for NRI -> too many features.

@@

```{r eda-10 }
# remove response
bands_data_corrected_buffer2_no_response <- bands_data_corrected_buffer2
bands_data_corrected_buffer2_no_response$defoliation <- NULL

pca_hr <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2_no_response,
  variance_cap = 0.90,
  geom_label_args = list(size = 3),
  ggtheme = ggpubr::theme_pubclean()
)
pca_hr[[1]]
```

@@

```{r buffer-size-comp-plot}
map(
  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
  ~ compare_buffer(
    trees_with_bands_corrected_buffer2,
    trees_with_bands_corrected_no_buffer,
    .x
  )
) %>%
  patchwork::wrap_plots() +
  patchwork::guide_area() +
  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
  patchwork::plot_annotation(
    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
    subtitle = "Mean values of each hyperspectral band for each plot "
  )
```

# PCA feature set comparison

Comparing the "proportion of variance explained" (POV) for all feature sets.

```{r pca-pov-comparison, dev = c("png", "pdf")}
pca_vi[[1]] / pca_nri[[1]] / pca_hr[[1]] *
  labs(title = NULL) *
  theme(
    axis.title.y = element_text(size = 8),
    axis.title.x = element_text(size = 8)
  ) +
  patchwork::plot_annotation(
    tag_levels = list(c("VI", "NRI", "HR")),
    title = "Proportin of variance explained by principal components",
    subtitle = "Labels indicate cumulative % of explained variance"
  )
```
